### 2024-04-02 paper/paper.qmd
paper

---
title: "Analyzing the Influence of Gender and Age on Shelter Return numbers in Toronto In 2023"
author: 
  - Xuecheng Gao
thanks: "https://github.com/XuechengGao/Shelter_Return_numbers_in_Toronto_2023.git"
date: today
date-format: long
abstract: "This study examines the impact of gender and age on Toronto shelter return numbers in 2023. We used linear regression to analyze data from shelter records to examine the relationship between gender, age, and shelter returns. The study found significant correlations between gender, age demographics and housing return rates. These results illustrate the importance of considering gender and age demographics in addressing homelessness in the Toronto region."
format: pdf
number-sections: true
bibliography: references.bib
---

```{r}
#| include: false
#| warning: false
#| message: false
#### Workspace setup ####
library(ggplot2)
library(arrow)
library(scales)
library(modelsummary)
library(tidyverse)
library(rstanarm)
library(knitr)
library(here)
library(marginaleffects)
```

# Introduction

# Data {#sec-data}

## Data Source

## Data clean

## Methodology

## Variables

# Model

## Model set-up

## Model Justification

# Results {#sec-results}

# Discussion {#sec-discussion}

## First discussion point {#sec-first-point}

## Second discussion point

## Third discussion point

## Weaknesses and next steps

\newpage

\appendix

# Appendix {-}

# Additional data details

# Model details {#sec-model-details}

\newpage

# References